---
title: "Grided space-filling design with hetGP surrogates for Fish"
author: "Jiangeng Huang"
date: "June 27, 2020"
output: html_document

params:
  # Set Netlogo software and model paths
  nl.path:  "~/fish/NetLogo 6.0.4/app"
  nl.jarname:  "netlogo-6.0.4.jar"
  model.path:  "/models/fish/Fish.nlogo"
  wd:  "~/fish/NetLogo 6.0.4/app/models/fish/"
  
  # True if you would like results written to file (results.csv)
  write_to_file: FALSE
  
  # Set initial paramaters for the fish model. Parameter values can be in the form of a single value or a list of values (e.g., population: [1000,2000])
  population_from_distribution: FALSE # True if the population size should be drawn from an exponential distribution
  population_mean: 1000 # If population_from_distribution is true, this is the mean of the exponential
  population:  [180, 200, 500, 1000] # If population_from_distribution is false, use this population size

  capture_line_length:  10 # the length of the capture line (net) in number of parcels
  use_capture_time:  FALSE # True if each capture session ends after a certain number of ticks; 
                           # False if each capture session ends after a certain number of fish are caught  
  num_to_capture:  100 # number of fish to mark and re-capture at each capture session
  capture_time_line:  90 # length of time (in number of ticks) that each capture session should last
  first_capture: 400 # length of time (in number of ticks) before first capture begins
  time_between_captures: 200 # length of time (in number of ticks) between capture sessions
  total_captures:  2 # total number of captures the simulation should run
  
  num_cores: 6 # number of cores to use for parallel processing
  runs: 20 # number of times to run simulation
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
unlockBinding("params", env = .GlobalEnv)

# libraries
Sys.setenv(NOAWT=1) # runs Rnetlogo in headless mode
library(RNetLogo)
library(dplyr)
library(hetGP)
library(stringr)
library(reshape2)
library(parallel)
library(lhs)
library("data.table")
options(warn=-1)
```

## Model Description

This program runs and outputs results of an agent-based model developed in NetLogo (Wilensky, 1999). The model simulates the dynamics of the capture, re-capture problem in animal populations. The environment is a 70 x 70 grid space. The fish agents move ("swim") on a grid space and are "captured" if they swim across a capture line (net). Fish are assumed to be immediately placed back into into the environments where they can potentially be re-captured. The user can select values for several parameters, including the size of the fish population, the length of the capture line, the number of fish to capture at a time, and the total captures to simulate.

This model builds on the netlogo Flocking model of birds (Wilensky, 1998). For more details on the model, please see the 'Info' tab in Fish.nlogo.

## Parameter Settings

#### Population Parameters
The population size can be defined or the simulation can draw the size of the population from an exponential distribution. 

Population size drawn from exponential distribution: `r params$population_from_distribution`  
Mean population size (if drawing from exponential distribution): `r params$population_mean`  
Population size (if not pulling from exponential distribution): `r params$population`  

#### Capture Parameters
A capture session can last either for a predefined number of ticks (capture time) or until a certain number of fish have been "caught" (number of fish to capture).  

Length of fish net (in parcels): `r params$capture_line_length`  
Time between captures (in ticks): `r params$time_between_captures`
Use capture time:  `r params$use_capture_time`  
Capture time (in ticks) (if use capture time is true):  `r params$capture_time_line`  
Number of fish to capture at each session (if use capture time is false):  `r params$num_to_capture` 
 
#### Simulation Parameters
The simulation will run into its completed all of the capture sessions.  

Number of capture sessions:  `r params$total_captures`  
Number of runs:  `r params$runs`  


```{r netlogo}
# function to run netlogo and place results in a dataframe
source(paste0(params$wd,"netlogo_fish_functions.R"))

## set up boundary for X values: 
lower_b <- 150
upper_b <- 4000

## maximinLHS: 
set.seed(2)
#x_ori <- maximinLHS(20, 1)      # maximin Latin square sampling
x_ori <- seq(0, 1, length.out = 20)  # grided design
params$population <- as.integer( lower_b + x_ori * (upper_b - lower_b) )  # round to nearest integer

# to ensure we're doing the correct number of runs, make any necessary parameter adjustments given certain criteria
if (params$population_from_distribution == TRUE) { params$population = 0 }
if (params$use_capture_time == TRUE) { params$num_to_capture == 0 } else { params$capture_time_line == 0 }

## create table with all parameter combinations: 
## the unique design points for X settings: 

parameters.df <- expand.grid(population_from_distribution = params$population_from_distribution, population_mean = params$population_mean,
                             population = params$population, capture_line_length = params$capture_line_length, 
                             num_to_capture = params$num_to_capture, capture_time_line = params$capture_time_line,
                             use_capture_time = params$use_capture_time, total_captures = params$total_captures,
                             time_between_captures = params$time_between_captures,
                             stringsAsFactors = F)
```


## Run fish model: 

```{r}
## Due to memory limits and cluster issues, run the code in 
## smaller batches in 'for' loop for multiple times. 
## Then, wrapped up the generated data sets together in the end. 

## Note: there can be occasionally  "checkForRemoteErrors(val)" issues in 
## the cluster when any of the fish model broke

# define number of batches 

n_batch  <- 25

# create a data list: 
data_all <- list()

for (j in 20:n_batch) {

ptm <- proc.time()
out <- run_netlogo(parameters.df, params$nl.path, params$nl.jarname, params$model.path, params$runs, params$num_cores)
sim_time <- proc.time() - ptm; sim_time

# count number of fish caught in second capture that were also caught in first capture
data1 <- filter(out,capture_num < 3)
data1 <- group_by(data1,who,runID,population_size) %>% summarise(captures = sum(captured))
data1 <- group_by(data1,runID,population_size,captures) %>% summarise(recaptures=n())
data1 <- filter(data1,captures==max(captures))
data1$recaptures <- ifelse(data1$captures<2,0,data1$recaptures)
data1 <- data1[c(1,2,4)]


## If missing: run it again 
if (  length(data1$population_size) == 0  ) {
  # run fish model at the new run
out <- run_netlogo(parameters.newrun, params$nl.path, params$nl.jarname, params$model.path, params$runs, params$num_cores)


# count number of fish caught in second capture that were also caught in first capture
data1 <- filter(out,capture_num < 3)
data1 <- group_by(data1,who,runID,population_size) %>% summarise(captures = sum(captured))
data1 <- group_by(data1,runID,population_size,captures) %>% summarise(recaptures=n())
data1 <- filter(data1,captures==max(captures))
data1$recaptures <- ifelse(data1$captures<2,0,data1$recaptures)
data1 <- data1[c(1,2,4)]
}

write.csv(data1,paste0(params$wd,"/out_summary.csv"), row.names = F)

## Save the simulated data: 
write.csv(data1, file = paste("GridData_3_", j, ".csv"),row.names=FALSE)

## Save a data frame in the environment
assign(paste("data_", j, sep = ""), data1)

data_all[[j]] <- data1

print(paste("Finished the batch number", j))
}
```

## Wrap up all the batches of runs: 

```{r}
data1  <- rbindlist(data_all)
write.csv(data1[, 2:3], file = paste("GridData_2.csv"),row.names=FALSE)
```


## Fit models: 

```{r}
## First hetGP fit: 
X_ori <- data1$population_size

## Scale X in [100, 2000] into [0 ,1] unit cube: at population=100, the recapture is 100% 
X <- ( data1$population_size - lower_b )/ (upper_b - lower_b)   # between 1 to 2000 since 0 never stops

## Take sqrt of Y: 
Y <- sqrt(data1$recaptures)

## Fit a hetGP: 
mod.a <- mleHetGP(X = X, Z = Y, lower = 0.0001, upper = 10)

## Fit a homGP: 
mod.b <- mleHomGP(X = X, Z = Y, lower = 0.0001, upper = 10)

## Fit a het.t process: 
mod.c <- mleHetTP(X = X, Z = Y, lower = 0.0001, upper = 10)
```

## One-shot Uniform design with hetGP visualization: 

```{r plot, echo=FALSE}
xgrid <- seq(0, 1, length = 1000)
p.a <- predict(mod.a, matrix(xgrid, ncol = 1))
pvar.a <- p.a$sd2 + p.a$nugs

p.b <- predict(mod.b, matrix(xgrid, ncol = 1))
pvar.b <- p.b$sd2 + p.b$nugs

p.c <- predict(mod.c, matrix(xgrid, ncol = 1))
pvar.c <- p.c$sd2 + p.c$nugs

## HetGP fit: 

## in sqrt scale for population: 
plot(xgrid* (upper_b - lower_b) + lower_b, p.a$mean, type = "l",xlab = "Population", ylab = "Sqrt of Number of Marked in Recapture",   main ="HetGP with Uniform Design (Sqrt Output)", ylim = c(0, 10))

points(X_ori, Y)
segments(mod.a$X0 * (upper_b - lower_b) + lower_b, rep(0, nrow(mod.a$X0)) , mod.a$X0 * (upper_b - lower_b) + lower_b, (mod.a$mult ) * 0.1,    col = "gray")
lines(xgrid * (upper_b - lower_b) + lower_b, qnorm(0.05, p.a$mean, sqrt(pvar.a)), col = 2, lty = 2)
lines(xgrid * (upper_b - lower_b) + lower_b, qnorm(0.95, p.a$mean, sqrt(pvar.a)), col = 2, lty = 2)

## in original scale for population: 
plot(xgrid*  (upper_b - lower_b) + lower_b, (p.a$mean)^2, type = "l", xlab = "Population", ylab = "Number of Marked in Recapture",    main = "HetGP with Uniform Design (Original Output)", ylim = c(0, 100))
points(X_ori, Y^2)
segments(mod.a$X0 *  (upper_b - lower_b) + lower_b, rep(0, nrow(mod.a$X0)) , mod.a$X0 *  (upper_b - lower_b) + lower_b, (mod.a$mult ) ,    col = "gray")
lines(xgrid *  (upper_b - lower_b) + lower_b, (qnorm(0.05, p.a$mean, sqrt(pvar.a)))^2, col = 2, lty = 2)
lines(xgrid *  (upper_b - lower_b) + lower_b, (qnorm(0.95, p.a$mean, sqrt(pvar.a)))^2, col = 2, lty = 2)

## Hom GP fit: 

## in sqrt scale for population: 
plot(xgrid* (upper_b - lower_b) + lower_b, p.b$mean, type = "l",xlab = "Population", ylab = "Sqrt of Number of Marked in Recapture",   main ="HomGP with Uniform Design (Sqrt Output)", ylim = c(0, 10))

points(X_ori, Y)
segments(mod.b$X0 * (upper_b - lower_b) + lower_b, rep(0, nrow(mod.b$X0)) , mod.b$X0 * (upper_b - lower_b) + lower_b, (mod.b$mult ) * 0.1,    col = "gray")
lines(xgrid * (upper_b - lower_b) + lower_b, qnorm(0.05, p.b$mean, sqrt(pvar.b)), col = 2, lty = 2)
lines(xgrid * (upper_b - lower_b) + lower_b, qnorm(0.95, p.b$mean, sqrt(pvar.b)), col = 2, lty = 2)

## in original scale for population: 
plot(xgrid*  (upper_b - lower_b) + lower_b, (p.b$mean)^2, type = "l", xlab = "Population", ylab = "Number of Marked in Recapture",    main = "HomGP with Uniform Design (Original Output)", ylim = c(0, 100))
points(X_ori, Y^2)
segments(mod.b$X0 *  (upper_b - lower_b) + lower_b, rep(0, nrow(mod.b$X0)) , mod.b$X0 *  (upper_b - lower_b) + lower_b, (mod.b$mult ) ,    col = "gray")
lines(xgrid *  (upper_b - lower_b) + lower_b, (qnorm(0.05, p.b$mean, sqrt(pvar.b)))^2, col = 2, lty = 2)
lines(xgrid *  (upper_b - lower_b) + lower_b, (qnorm(0.95, p.b$mean, sqrt(pvar.b)))^2, col = 2, lty = 2)

## HetTP fit: 

## in sqrt scale for population: 
plot(xgrid* (upper_b - lower_b) + lower_b, p.c$mean, type = "l",xlab = "Population", 
     ylab = "Sqrt of Number of Marked in Recapture",   main ="HetTP with Uniform Design (Sqrt Output)", ylim = c(0, 10))

points(X_ori, Y)
segments(mod.c$X0 * (upper_b - lower_b) + lower_b, rep(0, nrow(mod.c$X0)) , 
         mod.c$X0 * (upper_b - lower_b) + lower_b, (mod.c$mult ) * 0.1,    col = "gray")
lines(xgrid * (upper_b - lower_b) + lower_b, p.c$mean + 2 * sqrt(p.c$sd2 + p.c$nugs), col = 2, lty = 2)
lines(xgrid * (upper_b - lower_b) + lower_b, p.c$mean - 2 * sqrt(p.c$sd2 + p.c$nugs), col = 2, lty = 2)

## in original scale for population: 
plot(xgrid*  (upper_b - lower_b) + lower_b, (p.c$mean)^2, type = "l", xlab = "Population", 
     ylab = "Number of Marked in Recapture", main = "HetTP with Uniform Design (Original Output)", ylim = c(0, 100))
points(X_ori, Y^2)
segments(mod.c$X0 *  (upper_b - lower_b) + lower_b, rep(0, nrow(mod.c$X0)) , mod.c$X0 *  (upper_b - lower_b) + lower_b, (mod.c$mult ) ,    col = "gray")
lines(xgrid *  (upper_b - lower_b) + lower_b, (p.c$mean + 2 * sqrt(p.c$sd2 + p.c$nugs))^2, col = 2, lty = 2)
lines(xgrid *  (upper_b - lower_b) + lower_b, (p.c$mean - 2 * sqrt(p.c$sd2 + p.c$nugs))^2, col = 2, lty = 2)
```
