---
title: "Sequential design using hetGP for Fish"
author: "Jiangeng Huang"
date: "June 17, 2020"
output: html_document

params:
  # Set Netlogo software and model paths
  nl.path:  "~/fish/NetLogo 6.0.4/app"
  nl.jarname:  "netlogo-6.0.4.jar"
  model.path:  "/models/fish/Fish.nlogo"
  wd:  "~/fish/NetLogo 6.0.4/app/models/fish/"
  
  # True if you would like results written to file (results.csv)
  write_to_file: FALSE
  
  # Set initial paramaters for the fish model. Parameter values can be in the form of a single value or a list of values (e.g., population: [1000,2000])
  population_from_distribution: FALSE # True if the population size should be drawn from an exponential distribution
  population_mean: 1000 # If population_from_distribution is true, this is the mean of the exponential
  population:  [200, 500, 1000, 1500, 2000] # If population_from_distribution is false, use this population size
  capture_line_length:  10 # the length of the capture line (net) in number of parcels
  use_capture_time:  FALSE # True if each capture session ends after a certain number of ticks; 
                           # False if each capture session ends after a certain number of fish are caught  
  num_to_capture:  100 # number of fish to mark and re-capture at each capture session
  capture_time_line:  90 # length of time (in number of ticks) that each capture session should last
  first_capture: 400 # length of time (in number of ticks) before first capture begins
  time_between_captures: 200 # length of time (in number of ticks) between capture sessions
  total_captures:  2 # total number of captures the simulation should run
  
  num_cores: 6 # number of cores to use for parallel processing
  runs: 1 # number of times to run simulation
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
unlockBinding("params", env = .GlobalEnv)

# libraries
Sys.setenv(NOAWT=1) # runs Rnetlogo in headless mode
library(RNetLogo)
library(dplyr)
library(hetGP)
library(stringr)
library(reshape2)
library(parallel)
library(lhs)
options(warn=-1)
```

## Model Description

This program runs and outputs results of an agent-based model developed in NetLogo (Wilensky, 1999). The model simulates the dynamics of the capture, re-capture problem in animal populations. The environment is a 70 x 70 grid space. The fish agents move ("swim") on a grid space and are "captured" if they swim across a capture line (net). Fish are assumed to be immediately placed back into into the environments where they can potentially be re-captured. The user can select values for several parameters, including the size of the fish population, the length of the capture line, the number of fish to capture at a time, and the total captures to simulate.

This model builds on the netlogo Flocking model of birds (Wilensky, 1998). For more details on the model, please see the 'Info' tab in Fish.nlogo.

## Parameter Settings

#### Population Parameters
The population size can be defined or the simulation can draw the size of the population from an exponential distribution. 

Population size drawn from exponential distribution: `r params$population_from_distribution`  
Mean population size (if drawing from exponential distribution): `r params$population_mean`  
Population size (if not pulling from exponential distribution): `r params$population`  

#### Capture Parameters
A capture session can last either for a predefined number of ticks (capture time) or until a certain number of fish have been "caught" (number of fish to capture).  

Length of fish net (in parcels): `r params$capture_line_length`  
Time between captures (in ticks): `r params$time_between_captures`
Use capture time:  `r params$use_capture_time`  
Capture time (in ticks) (if use capture time is true):  `r params$capture_time_line`  
Number of fish to capture at each session (if use capture time is false):  `r params$num_to_capture` 
 
#### Simulation Parameters
The simulation will run into its completed all of the capture sessions.  

Number of capture sessions:  `r params$total_captures`  
Number of runs:  `r params$runs`  


```{r netlogo}
# function to run netlogo and place results in a dataframe
source(paste0(params$wd,"netlogo_fish_functions.R"))

## set up boundary for X values: 
lower_b <- 150
upper_b <- 4000

# to ensure we're doing the correct number of runs, make any necessary parameter adjustments given certain criteria
if (params$population_from_distribution == TRUE) { params$population = 0 }
if (params$use_capture_time == TRUE) { params$num_to_capture == 0 } else { params$capture_time_line == 0 }

## create table with all parameter combinations: 
## the unique design for X settings: 

## Initial design: 
set.seed(2)
#x_ori <- maximinLHS(10, 1)
x_ori <- seq(0, 1, length.out = 5)  # grided design
params$population <- as.integer( lower_b + x_ori * (upper_b - lower_b) )

parameters.df <- expand.grid(population_from_distribution = params$population_from_distribution, population_mean = params$population_mean,
                             population = params$population, capture_line_length = params$capture_line_length, 
                             num_to_capture = params$num_to_capture, capture_time_line = params$capture_time_line,
                             use_capture_time = params$use_capture_time, total_captures = params$total_captures,
                             time_between_captures = params$time_between_captures,
                             stringsAsFactors = F)

params$runs <- 5  # number of replications at initial design

# initial run of fish model
out <- run_netlogo(parameters.df, params$nl.path, params$nl.jarname, params$model.path, params$runs, params$num_cores)

# count number of fish caught in second capture that were also caught in first capture
data1 <- filter(out,capture_num < 3)
data1 <- group_by(data1,who,runID,population_size) %>% summarise(captures = sum(captured))
data1 <- group_by(data1,runID,population_size,captures) %>% summarise(recaptures=n())
data1 <- filter(data1,captures==max(captures))
data1$recaptures <- ifelse(data1$captures<2,0,data1$recaptures)
data1 <- data1[c(1,2,4)]

plot(data1$population_size, data1$recaptures, xlab="Population", ylab="Recaptured")

write.csv(data1,paste0(params$wd,"/out_summary.csv"), row.names = F)

## Initial hetGP fit: 
X_ori <- data1$population_size

## Scale X in [100, 2000] into [0 ,1] unit cube: at population=100, the recapture is 100% 
X <- ( data1$population_size - lower_b )/ (upper_b - lower_b)   

## save output
Y <- data1$recaptures

## Sequential design based on sqrt fit
mod.a <- mleHetGP(X = X, Z = sqrt(Y), lower = 0.0001, upper = 10)

## Total design budget: 
N <- 400

## Lookahead: 
h <- rep(NA, N - length(X))

## Set 1 run at each iteration
params$runs <- 1
   
## Sequential runs: 
 for(i in 1:( N - length(X))) {
   
   h[i] <- horizon(mod.a)   ## Adaptive horizon
   opt <- IMSPE_optim(mod.a, h = h[i])
   
   ## collect the new design: 
   X <- c(X, opt$par)
   
   ## run the new point: 
   
   ## set up for the model: 
parameters.newrun <- expand.grid(population_from_distribution = params$population_from_distribution, 
                                 population_mean = params$population_mean,
                                 population = as.integer(opt$par * (upper_b - lower_b) + lower_b),  ## change input population as integer
                                 capture_line_length = params$capture_line_length,
                                num_to_capture = params$num_to_capture, 
                                capture_time_line = params$capture_time_line,
                             use_capture_time = params$use_capture_time, total_captures = params$total_captures,
                             time_between_captures = params$time_between_captures,
                             stringsAsFactors = F)

# run fish model at the new run
out <- run_netlogo(parameters.newrun, params$nl.path, params$nl.jarname, params$model.path, params$runs, params$num_cores)

## Organize the new design and data: 
data_new <- filter(out, capture_num < 3)
data_new <- group_by(data_new, who, runID, population_size) %>%  summarise(captures = sum(captured))

data_new <- group_by(data_new,runID,population_size,captures) %>% summarise(recaptures=n())
data_new <- filter(data_new,captures==max(captures))
data_new$recaptures <- ifelse(data_new$captures<2,0,data_new$recaptures)
data_new <- data_new[c(1,2,4)]


## If missing: run it again 
if (  length(data_new$population_size) == 0  ) {
  # run fish model at the new run
out <- run_netlogo(parameters.newrun, params$nl.path, params$nl.jarname, params$model.path, params$runs, params$num_cores)

## Organize the new design and data: 
data_new <-filter(out, captured==TRUE) # filter for fish that were captured at least once
data_new <-filter(data_new, capture_num <= params$total_captures) # filter out any fish that may have been caught in extra capture at end of simulation
# count number of fish caught in second capture that were also caught in first capture
data_new <- filter(data_new, capture_num < 3)
data_new <- group_by(data_new, who, runID, population_size) %>% summarise(captures = n())

## TODO: This step can lead to empty data_new when no fish recaptured: 
data_new <- filter(data_new, captures == 2) %>% group_by(., runID, population_size) %>% summarise(recaptures=n())
}

## Record new design in integer:
X_ori  <- c(X_ori, data_new$population_size) # in original scale

## Collect new data: 

Ynew <- data_new$recaptures
Y <- c(Y, Ynew)

## update the model:
mod.a <- update(mod.a, Xnew = opt$par, Znew = sqrt(Ynew), ginit = mod.a$g * 1.01)

## Periodically restart MLE calculation: refresh
if(i %% 10 == 0){ mod2 <- mleHetGP(X = list(X0 = mod.a$X0, Z0 = mod.a$Z0, mult = mod.a$mult), Z = mod.a$Z, lower = 0.0001, upper = 1)
if(mod2$ll > mod.a$ll) mod.a <- mod2
}

print(i)
 }

data_seq <- data.frame(population_size = X_ori, recaptures = Y)

## Save sequential simulated data:
write.csv(data_seq, file = "SeqData.csv",row.names=FALSE)
```


### Fit model and make plots  

```{r}
## First hetGP fit: 
X_ori <- data_seq$population_size

## Scale X in [100, 2000] into [0 ,1] unit cube: at population=100, the recapture is 100% 
X <- ( data_seq$population_size - lower_b )/ (upper_b - lower_b)   # between 1 to 2000 since 0 never stops

## Take sqrt of Y: 
Y <- sqrt(data_seq$recaptures)

## Fit a hetGP: 
mod.a <- mleHetGP(X = X, Z = Y, lower = 0.0001, upper = 10)

## Fit a homGP: 
mod.b <- mleHomGP(X = X, Z = Y, lower = 0.0001, upper = 10)

## Fit a het.t process: 
mod.c <- mleHetTP(X = X, Z = Y, lower = 0.0001, upper = 10)
```

## One-shot Uniform design with hetGP visualization: 

```{r plot, echo=FALSE}
xgrid <- seq(0, 1, length = 1000)
p.a <- predict(mod.a, matrix(xgrid, ncol = 1))
pvar.a <- p.a$sd2 + p.a$nugs

p.b <- predict(mod.b, matrix(xgrid, ncol = 1))
pvar.b <- p.b$sd2 + p.b$nugs

p.c <- predict(mod.c, matrix(xgrid, ncol = 1))
pvar.c <- p.c$sd2 + p.c$nugs

## HetGP fit: 

## in sqrt scale for population: 
plot(xgrid* (upper_b - lower_b) + lower_b, p.a$mean, type = "l",xlab = "Population", ylab = "Sqrt of Number of Marked in Recapture",   main ="HetGP with Sequential Design (Sqrt Output)", ylim = c(0, 10))

points(X_ori, Y)
segments(mod.a$X0 * (upper_b - lower_b) + lower_b, rep(0, nrow(mod.a$X0)) , mod.a$X0 * (upper_b - lower_b) + lower_b, (mod.a$mult ) * 0.1,    col = "gray")
lines(xgrid * (upper_b - lower_b) + lower_b, qnorm(0.05, p.a$mean, sqrt(pvar.a)), col = 2, lty = 2)
lines(xgrid * (upper_b - lower_b) + lower_b, qnorm(0.95, p.a$mean, sqrt(pvar.a)), col = 2, lty = 2)

## in original scale for population: 
plot(xgrid*  (upper_b - lower_b) + lower_b, (p.a$mean)^2, type = "l", xlab = "Population", ylab = "Number of Marked in Recapture",    main = "HetGP with Sequential Design (Original Output)", ylim = c(0, 100))
points(X_ori, Y^2)
segments(mod.a$X0 *  (upper_b - lower_b) + lower_b, rep(0, nrow(mod.a$X0)) , mod.a$X0 *  (upper_b - lower_b) + lower_b, (mod.a$mult ) ,    col = "gray")
lines(xgrid *  (upper_b - lower_b) + lower_b, (qnorm(0.05, p.a$mean, sqrt(pvar.a)))^2, col = 2, lty = 2)
lines(xgrid *  (upper_b - lower_b) + lower_b, (qnorm(0.95, p.a$mean, sqrt(pvar.a)))^2, col = 2, lty = 2)

## Hom GP fit: 

## in sqrt scale for population: 
plot(xgrid* (upper_b - lower_b) + lower_b, p.b$mean, type = "l",xlab = "Population", ylab = "Sqrt of Number of Marked in Recapture",   main ="HomGP with Sequential Design (Sqrt Output)", ylim = c(0, 10))

points(X_ori, Y)
segments(mod.b$X0 * (upper_b - lower_b) + lower_b, rep(0, nrow(mod.b$X0)) , mod.b$X0 * (upper_b - lower_b) + lower_b, (mod.b$mult ) * 0.1,    col = "gray")
lines(xgrid * (upper_b - lower_b) + lower_b, qnorm(0.05, p.b$mean, sqrt(pvar.b)), col = 2, lty = 2)
lines(xgrid * (upper_b - lower_b) + lower_b, qnorm(0.95, p.b$mean, sqrt(pvar.b)), col = 2, lty = 2)

## in original scale for population: 
plot(xgrid*  (upper_b - lower_b) + lower_b, (p.b$mean)^2, type = "l", xlab = "Population", ylab = "Number of Marked in Recapture",    main = "HomGP with Sequential Design (Original Output)", ylim = c(0, 100))
points(X_ori, Y^2)
segments(mod.b$X0 *  (upper_b - lower_b) + lower_b, rep(0, nrow(mod.b$X0)) , mod.b$X0 *  (upper_b - lower_b) + lower_b, (mod.b$mult ) ,    col = "gray")
lines(xgrid *  (upper_b - lower_b) + lower_b, (qnorm(0.05, p.b$mean, sqrt(pvar.b)))^2, col = 2, lty = 2)
lines(xgrid *  (upper_b - lower_b) + lower_b, (qnorm(0.95, p.b$mean, sqrt(pvar.b)))^2, col = 2, lty = 2)

## HetTP fit: 

## in sqrt scale for population: 
plot(xgrid* (upper_b - lower_b) + lower_b, p.c$mean, type = "l",xlab = "Population", 
     ylab = "Sqrt of Number of Marked in Recapture",   main ="HetTP with Sequential Design (Sqrt Output)", ylim = c(0, 10))

points(X_ori, Y)
segments(mod.c$X0 * (upper_b - lower_b) + lower_b, rep(0, nrow(mod.c$X0)) , 
         mod.c$X0 * (upper_b - lower_b) + lower_b, (mod.c$mult ) * 0.1,    col = "gray")
lines(xgrid * (upper_b - lower_b) + lower_b, p.c$mean + 2 * sqrt(p.c$sd2 + p.c$nugs), col = 2, lty = 2)
lines(xgrid * (upper_b - lower_b) + lower_b, p.c$mean - 2 * sqrt(p.c$sd2 + p.c$nugs), col = 2, lty = 2)

## in original scale for population: 
plot(xgrid*  (upper_b - lower_b) + lower_b, (p.c$mean)^2, type = "l", xlab = "Population", 
     ylab = "Number of Marked in Recapture", main = "HetTP with Sequential Design (Original Output)", ylim = c(0, 100))
points(X_ori, Y^2)
segments(mod.c$X0 *  (upper_b - lower_b) + lower_b, rep(0, nrow(mod.c$X0)) , mod.c$X0 *  (upper_b - lower_b) + lower_b, (mod.c$mult ) ,    col = "gray")
lines(xgrid *  (upper_b - lower_b) + lower_b, (p.c$mean + 2 * sqrt(p.c$sd2 + p.c$nugs))^2, col = 2, lty = 2)
lines(xgrid *  (upper_b - lower_b) + lower_b, (p.c$mean - 2 * sqrt(p.c$sd2 + p.c$nugs))^2, col = 2, lty = 2)
```

